---
title: "Summary of results"
format:
  html:
    code-fold: show
    code-tools: true
    code-copy: true
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 2
editor: visual
---

# Working R environment

```{r}
library(tidyverse) 
```

# Needed data tables

```{r}
# last updated April 29, 2023
# save(euk_prok_counts, table_grazerate_wflp_wprok_weuk,
#      table_wcalcs, biov_calc, c_factor_biov,
#      pgC_ml_biov_bysample, pgC_ml_sizefrac_bysample,
#      carbon_consumed, file = "input-data/all-dfs-for-Tables.RData")

load("input-data/all-dfs-for-Tables.RData", verbose = TRUE)
```

# Table 1: counts and rates

**Table 1.** Grazing experiments conducted at the Mid-Cayman rise included 9 from Von Damm and 4 from Piccard. Out of 14 total grazing assays, 5 were conducted in Isobaric Gas Tight chambers at in situ pressure. Temperature reflects the highest recorded temperature at time of fluid collection. Prokaryotic cell concentrations were derived from descrete fixed samples from the same fluid, while eukaryotic cell concentrations reported below are dervied from the T0 grazing experiment time points. Absent grazing rates include those that had a negative slope and percentage prokaryote turnover shows the relative top-down (higher percentage) to bottom-up pressures on the microbial communities, based on grazing rate and cell concentrations.

```{r}
write_delim(table_wcalcs %>%
              select(SiteOrder, NameOrder, FLUID_ORIGIN, CRUISE_SAMPLE, EUK_ml, EUK_MinMax, PROK_ml, PROK_MinMax, GRAZING_EFFECT_hr, RATE_minute, CLEARANCE_RATE_ml, BAC_TURNOVER_PERC),
            file = "output-data/table1-grazing-exp-list.txt", delim = "\t")
```

## Average euk & prok summary

In results section, report the overall average euk cells per ml and prok cells per ml by sample type, location, experiment type, and vent field.

```{r, message=FALSE}
tmp <- euk_prok_counts %>% 
  mutate(TYPE = case_when(
    grepl("BSW", Name) ~ "non-vent",
    grepl("Plume", Name) ~ "non-vent",
    TRUE ~ "vent"
  ),
  EXP_TYPE = case_when(
    grepl("IGT", SITE_TYPE) ~ "IGT",
    TRUE ~ "Shipboard"
    ))

tmp %>% 
  select(TYPE, PROK_ml, EUK_ml) %>% 
  distinct() %>%
  group_by(TYPE) %>% 
  filter(!is.na(PROK_ml)) %>%
  filter(!is.na(EUK_ml)) %>%
  summarise(mean_prok = mean(PROK_ml),
            mean_euk = mean(EUK_ml))

tmp %>% 
  select(TYPE, PROK_ml, EUK_ml, Site) %>% 
  distinct() %>%
  group_by(TYPE, Site) %>% 
  filter(!is.na(PROK_ml)) %>%
  filter(!is.na(EUK_ml)) %>%
  summarise(mean_prok = mean(PROK_ml),
            mean_euk = mean(EUK_ml))

tmp %>% 
  select(TYPE, PROK_ml, EUK_ml, EXP_TYPE) %>% 
  distinct() %>%
  group_by(TYPE, EXP_TYPE) %>% 
  filter(!is.na(PROK_ml)) %>%
  filter(!is.na(EUK_ml)) %>%
  summarise(mean_prok = mean(PROK_ml),
            mean_euk = mean(EUK_ml))
```

## Complete grazing rate results

```{r}
table_wcalcs
```

Explanation of units for table with calculated values.

-   RATE_min & RATE_hr = Grazing rate as 'FLPs per grazer per minute' and per hour

-   CLEARANCE_RATE = ml or nL per grazer per hour

-   SPEC_GRAZE_RATE (Specific grazing rate) = Prokaryotes per grazer per hour

-   GRAZING EFFECT = bacteria per ml per hour

-   Bacterial turnover rate = % per day

## Average grazing rates

Report grazing rates from shipboard and IGT experiments. Differentiate vent and non-vent samples. Averages reported here do not include the undetected grazing rates.

```{r}
table_wcalcs %>% 
  # Technical replicates are already removed
  filter(!is.na(FLUID_ORIGIN)) %>% # Remove IGT 7 that didn't work
  filter(RATE_minute > 0) %>% # Remove undetected grazing rates
    mutate(SAMPLE_TYPE_BIN = case_when(
    NameOrder == "Background" ~ "Non-vent",
    NameOrder == "Plume" ~ "Non-vent",
    TRUE ~ "Vent"
  )) %>%
  # group_by(SAMPLE_TYPE_BIN) %>%
  separate(Experiment_rep, into = c("name", "exptype_rep"), remove = FALSE) %>% 
  mutate(exptype = case_when(exptype_rep == "Shipboard" ~ "Shipboard",
                             TRUE ~ "IGT")) %>% 
  group_by(SAMPLE_TYPE_BIN, exptype) %>%
  # group_by(SAMPLE_TYPE_BIN) %>% 
    summarise(MEAN_RATE = mean(GRAZING_EFFECT_hr),
              MIN_RATE = min(GRAZING_EFFECT_hr),
              MAX_RATE = max(GRAZING_EFFECT_hr))

table_wcalcs %>% 
  # Technical replicates are already removed
  filter(!is.na(FLUID_ORIGIN)) %>% # Remove IGT 7 that didn't work
  filter(RATE_minute > 0) %>% # Remove undetected grazing rates
    mutate(SAMPLE_TYPE_BIN = case_when(
    NameOrder == "Background" ~ "Non-vent",
    NameOrder == "Plume" ~ "Non-vent",
    TRUE ~ "Vent"
  )) %>%
  # group_by(SAMPLE_TYPE_BIN) %>%
  separate(Experiment_rep, into = c("name", "exptype_rep"), remove = FALSE) %>% 
  mutate(exptype = case_when(exptype_rep == "Shipboard" ~ "Shipboard",
                             TRUE ~ "IGT")) %>% 
  group_by(SAMPLE_TYPE_BIN, FIELD) %>% 
    summarise(MEAN_RATE = mean(GRAZING_EFFECT_hr),
              MIN_RATE = min(GRAZING_EFFECT_hr),
              MAX_RATE = max(GRAZING_EFFECT_hr))

# igt_vent_rate_avg <- 11400.71904
# ship_vent_rate_avg <- 2430.01651
# ship_nonvent_rate_avg <- 65.11676
```

Units above are in cells consumed per ml per hour.


Compare IGT and shipboard outputs with other studies below. Units: pg C per ml per hour

# Table 2: biomass

## Mean biovolume from microscopy images

Units are µm\^3 for biovolume.

```{r}
# Biovolume by sample type
# biov_calc
biov_calc %>% 
  group_by(VENT_BSW) %>% 
  summarise(mean_biov = mean(BIOVOLUME))

# Biovolume by IGT and sample type
biov_calc %>% 
  # filter(VENT_BSW != "BSW") %>% 
  group_by(EXP, VENT_BSW) %>% 
  summarise(mean_biov = mean(BIOVOLUME))

# write.csv(biov_calc, file = "output-data/biovolume-calculation.csv")
```

In discussion compare deep sea biovolumes to Pernice et al.
```{r}
biov_calc %>% 
  filter(VENT_BSW != "vent")
```


## Determine C conversion rate

Determined from biovolume, using pg of C per cell.

```{r}
pgC_cell_BSW <- subset(c_factor_biov, VARIABLES == "pgC_cell_MD_nodiatom")$mean_BSW

pgC_cell_vent <- subset(c_factor_biov, VARIABLES == "pgC_cell_MD_nodiatom")$mean_vent

# pg C per cell (pgC_cell)
pgC_cell_BSW # nonvent
pgC_cell_vent # vent
```

## Report carbon biomass as a mean

To estimate pg C per ml, multiply pg C per cell by euk cell concentration.

\*\*(cells/ml)\*(pg C/cell) = pg C/ml\*\*

Take the mean of estimate euk cells biomass across all samples (T0-Tfs). Remove the technical replicates first and convert pg C per ml to per L.

## Part 1 of Table 2

```{r}
## Vent vs. non vent
partONE <- pgC_ml_biov_bysample %>% 
  ##remove technical replicates:
  filter(IGT_REP != "IGT4b" & IGT_REP != "IGT5b") %>% 
  ##
  group_by(SAMPLE_TYPE_BIN)%>% 
  summarise(MEAN_pg_C_ml = round(mean(pgC_ml_mean), digits = 1),
            MAX_pg_C_ml = round(max(pgC_ml_mean), digits =1),
            MIN_pg_C_ml = round(min(pgC_ml_mean), digits = 1)) %>% 
  add_column(ORIGIN = "bybiovolume") %>% 
  select(CATEGORY = SAMPLE_TYPE_BIN, everything()) %>% 
  ## Von Damm vs Piccard
  rbind(
    pgC_ml_biov_bysample %>% 
  filter(IGT_REP != "IGT4b" & IGT_REP != "IGT5b") %>% 
  filter(SAMPLE_TYPE_BIN == "Vent") %>%
  group_by(Site)%>% 
  summarise(MEAN_pg_C_ml = round(mean(pgC_ml_mean), digits = 1),
            MAX_pg_C_ml = round(max(pgC_ml_mean), digits =1),
            MIN_pg_C_ml = round(min(pgC_ml_mean), digits = 1)) %>% 
  add_column(ORIGIN = "bybiovolume") %>% 
    select(CATEGORY = Site, everything())
  ) %>% 
  rbind(
    ## Vent only, IGT vs shipboard
    pgC_ml_biov_bysample %>% 
  filter(SAMPLE_TYPE_BIN == "Vent") %>%
  group_by(VARIABLE, EXP_TYPE, SAMPLE_TYPE_BIN) %>% 
  summarise(MEAN_pg_C_ml = round(mean(pgC_ml_mean), digits = 1),
            MAX_pg_C_ml = round(max(pgC_ml_mean), digits =1),
            MIN_pg_C_ml = round(min(pgC_ml_mean), digits = 1)) %>% 
    unite(CATEGORY, EXP_TYPE, VARIABLE, SAMPLE_TYPE_BIN, sep = " ") %>% 
  add_column(ORIGIN = "bybiovolume")) %>% 
  unite("MinMax_pg_C_ml", MIN_pg_C_ml, MAX_pg_C_ml, sep = " / ", remove = FALSE)
```

## Part 2 of Table 2

```{r}
## Vent vs. non vent
partTWO <- pgC_ml_sizefrac_bysample %>% 
  ##remove technical replicates:
  filter(IGT_REP != "IGT4b" & IGT_REP != "IGT5b") %>% 
  filter(VARIABLE == "totaleuk") %>% 
  ##
  group_by(SAMPLE_TYPE_BIN)%>% 
  summarise(MEAN_pg_C_ml = round(mean(pg_C_ml), digits = 1),
            MAX_pg_C_ml = round(max(pg_C_ml), digits =1),
            MIN_pg_C_ml = round(min(pg_C_ml), digits = 1)) %>% 
  add_column(ORIGIN = "sizefrac") %>% 
  select(CATEGORY = SAMPLE_TYPE_BIN, everything()) %>% 
  ## Von Damm vs Piccard
  rbind(
    pgC_ml_sizefrac_bysample %>% 
  filter(IGT_REP != "IGT4b" & IGT_REP != "IGT5b") %>% 
  filter(SAMPLE_TYPE_BIN == "Vent") %>%
    filter(VARIABLE == "totaleuk") %>% 
  group_by(Site)%>% 
  summarise(MEAN_pg_C_ml = round(mean(pg_C_ml), digits = 1),
            MAX_pg_C_ml = round(max(pg_C_ml), digits =1),
            MIN_pg_C_ml = round(min(pg_C_ml), digits = 1)) %>% 
  add_column(ORIGIN = "sizefrac") %>% 
    select(CATEGORY = Site, everything())
  ) %>% 
  rbind(
    ## Vent only, IGT vs shipboard
    pgC_ml_sizefrac_bysample %>% 
  filter(SAMPLE_TYPE_BIN == "Vent") %>%
    filter(VARIABLE == "totaleuk") %>% 
  group_by(VARIABLE, EXP_TYPE, SAMPLE_TYPE_BIN) %>% 
  summarise(MEAN_pg_C_ml = round(mean(pg_C_ml), digits = 1),
            MAX_pg_C_ml = round(max(pg_C_ml), digits =1),
            MIN_pg_C_ml = round(min(pg_C_ml), digits = 1)) %>% 
    unite(CATEGORY, EXP_TYPE, VARIABLE, SAMPLE_TYPE_BIN, sep = " ") %>% 
  add_column(ORIGIN = "sizefrac")) %>% 
  unite("MinMax_pg_C_ml", MIN_pg_C_ml, MAX_pg_C_ml, sep = " / ", remove = FALSE)
```

### Table 2

```{r}
table2 <- partONE %>% 
  rbind(partTWO) %>% 
  select(-starts_with("MinMax")) %>% 
  pivot_longer(cols = ends_with("C_ml"), values_to = "pg_C_ml") %>% 
  mutate(µg_C_L = (pg_C_ml*(0.000001)*1000)) %>% 
  mutate(STAT = str_remove(name, "_pg_C_ml")) %>% 
  select(-name) %>% 
  mutate(CLASSIFICATION = case_when(
    CATEGORY == "VD" ~ "Von Damm (vent only)",
    CATEGORY == "Piccard" ~ "Piccard (vent only)",
    grepl("Bag", CATEGORY) ~ "Shipboard (vent only)",
    grepl("IGT", CATEGORY) ~ "IGT (vent only)",
    TRUE ~ CATEGORY))

table2_ugC_L <- table2 %>% 
  select(CLASSIFICATION, µg_C_L, STAT, ORIGIN) %>% 
  pivot_wider(names_from = STAT, values_from = µg_C_L)

# write.csv(table2_ugC_L,
# file = "output-data/table2-BIOMASS_ugC_L.csv")
```


## Carbon consumed

Summary of carbon consumed:

```{r}
head(carbon_consumed)
```

```{r}
table3 <- carbon_consumed %>% 
  drop_na() %>% # Remove extra IGT7
  filter(GRAZING_EFFECT_hr > 0) %>% # Remove slope 0 experiments
  mutate(SAMPLE_TYPE_BIN = case_when(
    NameOrder == "Background" ~ "Non-vent",
    NameOrder == "Plume" ~ "Non-vent",
    TRUE ~ "Vent"
  )) %>%
  group_by(SAMPLE_TYPE_BIN) %>% 
  summarise(MEAN_pgC_pergrazer_CLEARANCE = round(mean(pgC_pergrazer_CLEARANCE), digits = 1),
            MEAN_pgC_perml_perhr = round(mean(pgC_perml_perhr), digits = 1),
            MAX_pgC_pergrazer_CLEARANCE = round(max(pgC_pergrazer_CLEARANCE), digits = 1),
            MAX_pgC_perml_perhr = round(max(pgC_perml_perhr), digits = 1),
            MIN_pgC_pergrazer_CLEARANCE = round(min(pgC_pergrazer_CLEARANCE), digits = 1),
            MIN_pgC_perml_perhr = round(min(pgC_perml_perhr), digits = 1)) %>% 
  select(CATEGORY = SAMPLE_TYPE_BIN, everything()) %>% 
  rbind(carbon_consumed %>% 
  drop_na() %>% # Remove extra IGT7
    filter(GRAZING_EFFECT_hr > 0) %>% # Remove slope 0 experiments
  mutate(SAMPLE_TYPE_BIN = case_when(
    NameOrder == "Background" ~ "Non-vent",
    NameOrder == "Plume" ~ "Non-vent",
    TRUE ~ "Vent"
  )) %>%
  filter(SAMPLE_TYPE_BIN == "Vent") %>% 
  group_by(SiteOrder) %>% 
    summarise(MEAN_pgC_pergrazer_CLEARANCE = round(mean(pgC_pergrazer_CLEARANCE), digits = 1),
            MEAN_pgC_perml_perhr = round(mean(pgC_perml_perhr), digits = 1),
            MAX_pgC_pergrazer_CLEARANCE = round(max(pgC_pergrazer_CLEARANCE), digits = 1),
            MAX_pgC_perml_perhr = round(max(pgC_perml_perhr), digits = 1),
            MIN_pgC_pergrazer_CLEARANCE = round(min(pgC_pergrazer_CLEARANCE), digits = 1),
            MIN_pgC_perml_perhr = round(min(pgC_perml_perhr), digits = 1)) %>% 
  select(CATEGORY = SiteOrder, everything())) %>% 
  rbind(carbon_consumed %>% 
  drop_na() %>% # Remove extra IGT7
    filter(GRAZING_EFFECT_hr > 0) %>% # Remove slope 0 experiments
  mutate(SAMPLE_TYPE_BIN = case_when(
    NameOrder == "Background" ~ "Non-vent",
    NameOrder == "Plume" ~ "Non-vent",
    TRUE ~ "Vent"
  )) %>%
    mutate(EXP_TYPE = case_when(
    grepl("IGT", CRUISE_SAMPLE) ~ "IGT",
    TRUE ~ "Shipboard"
  )) %>%
  filter(SAMPLE_TYPE_BIN == "Vent") %>% 
  group_by(EXP_TYPE) %>% 
    summarise(MEAN_pgC_pergrazer_CLEARANCE = round(mean(pgC_pergrazer_CLEARANCE), digits = 2),
            MEAN_pgC_perml_perhr = round(mean(pgC_perml_perhr), digits = 1),
            MAX_pgC_pergrazer_CLEARANCE = round(max(pgC_pergrazer_CLEARANCE), digits = 2),
            MAX_pgC_perml_perhr = round(max(pgC_perml_perhr), digits = 1),
            MIN_pgC_pergrazer_CLEARANCE = round(min(pgC_pergrazer_CLEARANCE), digits = 2),
            MIN_pgC_perml_perhr = round(min(pgC_perml_perhr), digits = 1)) %>% 
  select(CATEGORY = EXP_TYPE, everything())) %>% 
  ## Modify table
  unite("MinMax_pgC_pergrazer_CLEARANCE", MIN_pgC_pergrazer_CLEARANCE, MAX_pgC_pergrazer_CLEARANCE, sep = " / ", remove = TRUE) %>% 
  unite("MinMax_pgC_perml_perhr", MIN_pgC_perml_perhr, MAX_pgC_perml_perhr, sep = " / ", remove = TRUE) %>% 
  mutate(CLASSIFICATION = case_when(
    CATEGORY == "Von Damm" ~ "Von Damm (vent only)",
    CATEGORY == "Piccard" ~ "Piccard (vent only)",
    grepl("Bag", CATEGORY) ~ "Shipboard (vent only)",
    grepl("IGT", CATEGORY) ~ "IGT (vent only)",
    TRUE ~ CATEGORY)) %>% 
  select(-CATEGORY)
```

### Table 3

```{r}
write.csv(table3, file = "output-data/table3-C-consumption.csv")
```


## Carbon consumed total and by cell?

Assuming 86 fg C per cell for prokaryotic cells. 1 fg == 0.001 pg

```{r}
## From above:
igt_vent_rate_avg <- 11400.71904
ship_vent_rate_avg <- 2430.01651
ship_nonvent_rate_avg <- 65.11676


mor_prok_c <- 86 # in fg C per cell

# Non vent cells consumed per ml per
(mor_prok_c)*(0.001) * (65.11676) # pg

# IGT
igt <- (mor_prok_c)*(0.001) * (11400.71904);igt

# Shipboard
ship <- (mor_prok_c)*(0.001) * (2430.01651);ship

# (igt+ship)/2
```


# Table S1

Import from Hu et al. (*Molecular Ecology*).

```{r}
env_tmp <- read.delim("../../microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/samplelist-metadata.txt")
# head(env_tmp)

mcr_metadata <- env_tmp %>% 
  filter(SITE == "VonDamm" | SITE  == "Piccard") %>% 
  mutate(FIELD =  case_when(
    SITE == "VonDamm"  ~ "Von Damm",
    SITE == "Piccard" ~ "Piccard")) %>% 
  filter(SAMPLETYPE != "Incubation") %>% 
  separate(SAMPLE, into = c("num", "loc", "loc2", "fluid_origin", "name", "FLUID_ORIGIN", "else"), remove = FALSE) %>% 
  select(-Sample_or_Control, -SAMPLEID, -ref_num, -SITE, SEQ_SAMPLE = SAMPLE,FLUID_ORIGIN, everything(), -num, -loc, -loc2, -fluid_origin, -name, -`else`)


```

Generate supplementary table.

```{r}
supp_table_env_mcr <- mcr_metadata %>% 
  select(FIELD, VENT, SAMPLETYPE, DEPTH, TEMP = temp, PH = pH, PERCSEA = percseawater, MG = mg, H2 = h2, H2S = h2s, CH4 = ch4, MICROBIAL = ProkConc, everything())

# Save supplementary metadata table:
# write.csv(supp_table_env_mcr, file = "output-data/supp_table_MCR_metadata.csv")
```

# Table S2

```{r}
# # ## Table S2
# write_delim(table_wcalcs %>%
#               select(SiteOrder, NameOrder, FLUID_ORIGIN, CRUISE_SAMPLE,
#                      TimePoints, starts_with("EXP"), starts_with("CTRL"), EUK_ml, EUK_MinMax, EUK_sem, PROK_ml, PROK_MinMax, PROK_sem, FLP_ml, GRAZE_RATE_DAY, Slope = RATE, RATE_minute, GRAZING_EFFECT_hr, CLEARANCE_RATE_ml, BAC_TURNOVER_PERC),
#             file = "output-data/tableS2-grazing-exp-list.txt", delim = "\t")
```

# Table S3

Calculations not generated in R.

# Table S4

```{r}
biov_calc
```



# Table S5
Report carbon biomass values, but include size fractions.

```{r}
table5_cbiomass_total <- pgC_ml_biov_bysample %>% 
  group_by(VARIABLE, Site, EXP_TYPE, SAMPLE_TYPE_BIN) %>% 
  summarise(MEAN_pg_C_L = mean(pgC_ml_mean/1000),
            MAX_pg_C_L = max(pgC_ml_mean/1000),
            MIN_pg_C_L = min(pgC_ml_mean/1000)) %>% 
  add_column(ORIGIN = "bybiovolume") %>% 
  rbind(pgC_ml_sizefrac_bysample %>% 
  group_by(VARIABLE, Site, SAMPLE_TYPE_BIN, EXP_TYPE) %>% 
  summarise(MEAN_pg_C_L = mean(pg_C_ml/1000),
            MAX_pg_C_L = max(pg_C_ml/1000),
            MIN_pg_C_L = min(pg_C_ml/1000)) %>% 
  add_column(ORIGIN = "sizefrac")
  )
```


# Figure 3

Estimates to create figure 3, a day at a food web at a deep sea vent site.
Use units µg C per L per day.

From these tables:
  euk_prok_counts
  table_grazerate_wflp_wprok_weuk
  table_wcalcs
  biov_calc
  c_factor_biov
  pgC_ml_biov_bysample
  pgC_ml_sizefrac_bysample
  carbon_consumed
  

Isolate vent-only information and remove the zero values.

### Chemosynthetic bacteria & archaea
Prokaryote pool first
```{r}
head(euk_prok_counts)
prok_fgC_cell <- 86 # fg C per cell

proks_only <- euk_prok_counts %>% 
  select(PROK_ml, Name, SITE_TYPE, Experiment) %>% 
  distinct() %>% 
  drop_na() %>% 
  filter(Name != "Plume" & Name != "BSW")
proks_only

mean_prok <- mean(proks_only$PROK_ml)
max_prok <- max(proks_only$PROK_ml)
min_prok <- min(proks_only$PROK_ml)

# Cells per ml
# fgC/cell * cell/ml * (1e-9 ug/1fg) * (1000 ml/1 L)
prok_fgC_cell * mean_prok * (1e-9) * 1000

prok_fgC_cell * max_prok * (1e-9) * 1000

prok_fgC_cell * min_prok * (1e-9) * 1000
```

ug C per L

### Carbon consumed by heterotrophic protists

Use input from Table 3. Focus on grazing rate, instead of clearance rate.

Grazing rate is in pg C per ml per hr.
```{r}
table3
```

```{r}
tmp <- table3 %>% 
  filter(CLASSIFICATION == "Vent") %>% 
  select(ends_with("_perhr")) %>% 
  separate(MinMax_pgC_perml_perhr, into = c("min", "max"), sep = " / ")


# Units are in pg C per ml per hour
## pgC/ml hr * (1000ml/1L) * (24 hr/1 day) * (1e-6 ug / 1pg)
(tmp$MEAN_pgC_perml_perhr) * (1000 * 24 * (1e-6))

as.numeric(tmp$min) * (1000 * 24 * (1e-6))

as.numeric(tmp$max) * (1000 * 24 * (1e-6))
```

### Microeuk carbon pool

```{r}
table2 %>% 
  filter(CLASSIFICATION == "Vent") %>% 
  select(STAT, µg_C_L, CLASSIFICATION, ORIGIN)


```



# Supplemental calulcations for within range

## Compare to Connell

From Conne/l et al., table 4. Heterotrophic bacteria, Daily carbon grazed - based on standing stock and mortality rate in Southern California.

min: 1.74 max: 28.8

units: µg C per L per day

To compare with units:ug C per ml per hour (this study)

```{r}
min <- 1.74
max <- 28.8

# ug to pg: 1ug = 1E6 pg
((min)/1000)/24

((max)/1000)/24

```


## Compare to Pernice et al.

Convert to ug C per L.

1 pg = 1e-6 ug
```{r}
shallow_pgC_ml <- 280
deep_pgC_ml <- 50

# pgC/ml * (1e-6 ug / 1 pg) * (1000 ml / 1)

shallow_pgC_ml * (1e-6) * 1000
deep_pgC_ml * (1e-6) * 1000
```


## Compare to Medina et al.

Review, so this includes several studies. Grazing effect in Table 3 is reported in bacteria per ml per day.

Min and max are both from Pachiadaki et al. deep water and at an interface. min: 18.7 max: 13600

Min and max in this study is 24 and 17285 cells consumed per ml per hour
